
```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "##",
  tidy = FALSE,
  error = FALSE,
  fig.width = 8,
  fig.height = 8)
options(width = 90)
options(max.print = 200)
```

# The CRAN database

[![Linux Build Status](https://travis-ci.org/metacran/crandb.svg?branch=master)](https://travis-ci.org/metacran/crandb)
[![Windows Build status](https://ci.appveyor.com/api/projects/status/github/metacran/crandb?svg=true)](https://ci.appveyor.com/project/gaborcsardi/crandb)

The CRAN database provides an API for programatically accessing all
meta-data of CRAN R packages. This API can be used for various purposes,
here are three examples I am woking on right now:
* Writing a package manager for R. The package manager can use the
  CRAN DB API to query dependencies, or other meta data.
* Building a search engine for CRAN packages. The DB itself does not
  provide a search API, but it can be (easily) mirrored in a search
  engine.
* Creating an RSS feed for the new, updated or archived packages on CRAN.

**Note that `crandb` is _NOT_ an official CRAN project, and is not supported
by CRAN.**

## The `crandb` API

### Packages

`package()` returns the latest version of a package:

```{r}
library(crandb)
package("dotenv")
```

A given version can be queried as well:

```{r}
package("httr", version = "0.3")
```

Or all versions:

```{r}
package("httr", version = "all")
```

### List of all packages

`list_packages()` lists all packages, in various formats, potentially including
archived packages as well:

```{r}
list_packages(from = "cranlogs", limit = 10, archived = FALSE)
```

### CRAN events

`events()` lists CRAN events, starting from the latest ones. New package
releases and archival can both be included in the list. By default the last
10 events are included:

```{r}
events()
```

### R and CRAN releases

The `releases()` function lists recent R versions, with their release dates.

```{r}
releases()
```

The CRAN packages that were current at the time of an R releases can be listed
with the `cran_releases()` function:

```{r}
cran_releases(version = "2.13.1")
```

## The raw API

This is the raw JSON API.

We will use the
[`httr`](https://github.com/hadley/httr) package to query it, and the
[`jsonlite`](https://github.com/jeroenooms/jsonlite) package to nicely
format it. [`magrittr`](https://github.com/smbache/magrittr) is loaded
for the pipes. We use `DB` to query the API and format the result.

```{r}
library(magrittr)
skip_lines <- function(text, head = 1e6, tail = 1e6) {
	text <- strsplit(text, "\n")[[1]]
	tail <- min(tail, max(0, length(text) - head))
	skip_text <- if (length(text) > head + tail) {
		paste("\n... not showing", length(text) - head - tail, "lines ...\n")
	} else {
		character()
	}
    c(head(text, head), skip_text, tail(text, tail)) %>%
		paste(collapse = "\n")
}
DB <- function(api, head = 1e6, tail = head) {
  paste0("http://crandb.r-pkg.org", "/", api) %>%
    httr::GET() %>%
	httr::content(as = "text", encoding = "UTF-8") %>%
	jsonlite::prettify() %>%
	skip_lines(head = head, tail = tail) %>%
	cat()
}
```

### `/:pkg` Latest version of a package

The result includes all fields verbatim from the DESCRIPTION file,
plus some extra:
* `date`: The date and time when the package was published on CRAN.
  This is needed, because especially old packages might not have some
  (or any) of the `Date`, `Date/Publication` or `Packaged` fields.
* `releases`: The R version(s) that were released when this version
  of the package was the latest on CRAN.
* The `Suggests`, `Depends`, etc. fields are formatted as named lists.

```{r}
DB("/magrittr")
```

### `/:pkg/:version` A specific version of a package

The format is the same as for the latest version.

```{r}
DB("/magrittr/1.0.0")
```

### `/:pkg/all` All versions of a package

The result is a list of package versions in the `versions` field, in the
format above, plus some extra:
* `name`: The name of the package.
* `title`: The title field of the package.
* `latest`: The latest version of the package.
* `archived`: Whether the package was archived.
* `timeline`: All versions and their release dates of the package.

```{r}
DB("/magrittr/all")
```

### `/-/all` All packages, in alphabetical order

Note that this API point _really_ returns a list of all active CRAN packages
(currently about 6,000), so it is a good idea to query the date in
chunks. `limit` specifies the number of records to return, and `startkey`
can be used to specify the first key to list. Note that the result will
include the full records of the package, all package versions.

```{r}
DB("/-/all?limit=3", head = 20)
```

### `/-/latest` Latest versions of all packages

This is similar to `/-/all`, but only the latest versions of
the packages are returned, instead of the complete records with
all versions.

```{r}
DB("/-/latest?limit=3", head = 20)
```

### `/-/desc` Short description of latest package versions

Latest versions of all packages, in alphabetical order. It also
contains the `title` fields of the packages. Only active (not archived)
packages are included.

```{r}
DB("/-/desc?limit=5")
```

### `/-/allall` Complete records for all packages, including archived ones

This is similar to `/-/all`, but lists the archived packages as well.

```{r}
DB("/-/desc?limit=2", head = 20)
```

### `/-/pkgreleases` Package releases

All versions of all packages, in the order of their release. Note that
this includes each version of each package separately, so it is a very
long list, and it is a good idea to use the `limit` parameter. `descending`
can be used to reverse the ordering, and start with the most recent
releases.

```{r}
DB("/-/pkgreleases?limit=3&descending=true", head = 20)
```

### `/-/archivals` Package archivals

Package archival events, sorted by date times. The latest version
of the package record is also included. Again, use the `limit` parameter
to query in chunks, and the `descending` parameter to reverse the order,
and see most recent archivals last.

```{r}
DB("/-/archivals?limit=3&descending=true", head = 20)
```

### `/-/events` Release and archival events

The union of `/-/pkgreleases` and `/-/archivals`, so it includes
all package releases and archivals.

```{r}
DB("/-/events?limit=3&descending=true", head = 20)
```

### `/-/releases` List of R versions

List of all R versions supported by the database. Currently this goes
back to version 2.0.0, older versions will be potentially added later.

```{r}
DB("/-/releases", head = 20)
```

### `/-/releasepkgs/:version` Packages that were current at an R release

All package records that were current when a given version of R
was released. This is essentially a snapshot of CRAN, for a given R
release.

```{r}
DB("/-/releasepkgs/2.15.3", head = 20)
```

### `/-/release/:version` Package versions that were current at an R release

Similar to the previous list, but it only includes the version numbers,
and not the records of the packages.

```{r}
DB("/-/release/2.15.3", head = 20)
```

### `/-/releasedesc/:version` Short description of CRAN snapshots

Similar to `/-/release`, but it also include the the `title` fields of
the packages.

```{r}
DB("/-/releasedesc/2.15.3", head = 20)
```

### `/-/topdeps/:version` Packages most depended upon

Top twenty packages. It includes all forms of dependencies, and it can be
restricted to dependencies that were in place at a given R release. 

```{r}
DB("/-/topdeps/3.1.1")
```

For the latest versions of the packages and their dependencies, you can set
`:version` to `dev`:

```{r}
DB("/-/topdeps/devel")
```

### `/-/deps/:version` Number of reverse dependencies

For all packages, not just the top twenty.

```{r}
DB("/-/deps/2.15.1", head = 20)
```
